\thetitle

### 2010-04-08 ITP10/hybrid.tex
final ? updates

\documentclass[runningheads]{llncs}
\ifx\pdftexversion\undefined
  \usepackage[dvips]{graphics}
\else
  \usepackage[pdftex]{graphics}
\fi
\usepackage{amsmath}
%include format.tex
%\usepackage{rotating}
\usepackage{graphicx}
\input{diagrams}
\newcommand{\thetitle}{Automated Machine-Checked Hybrid System Safety Proofs}
\newcommand{\eelis}{Eelis van der Weegen$^{\dagger}$}
\newcommand{\herman}{Herman Geuvers$^{\dagger\ddagger}$}
\newcommand{\dan}{Dan Synek$^{\dagger}$}
\newcommand{\adam}{Adam Koprowski$^{\P}$}

@@

\def\phi{\varphi}
\newcommand{\weg}[1]{}

\newcommand{\DN}{{\sf DN}\,}
\newcommand{\IR}{{\mathbf R}}
\newcommand{\Heat}{{\sf Heat}}
\newcommand{\Cool}{{\sf Cool}}
\newcommand{\Chec}{{\sf Check}}
\newcommand{\Off}{{\rm Off}} 
\newcommand{\Loc}{{\rm Loc}}
\newcommand{\AState}{{\rm AState}}
\newcommand{\State}{{\rm State}}
\newcommand{\Inv}{{\rm Inv}}
\newcommand{\Init}{{\rm Init}}
\newcommand{\Region}{{\rm Region}}
\newcommand{\List}{{\rm List}}
\newcommand{\Unsafe}{{\rm Unsafe}}
\newcommand{\Safe}{{\rm Safe}}
\newcommand{\AUnsafe}{{\rm AUnsafe}}
\newcommand{\ASafe}{{\rm ASafe}}
\newcommand{\Reach}{{\rm Reach}}
\newcommand{\AReach}{{\rm AReach}}
\newcommand{\guard}{{\rm guard}}
\newcommand{\reset}{{\rm reset}}
\newcommand{\trans}{\rightarrow_{CD}}
\newcommand{\doubleheadrightarrow}{\mathrel{\rightarrow\!\!\!\!\rightarrow}}
\newcommand{\ttrans}{\doubleheadrightarrow_{CD}}
\newcommand{\attrans}{\stackrel{A}{\doubleheadrightarrow_{CD}}}
\newcommand{\contrans}{\rightarrow_C}
\newcommand{\distrans}{\rightarrow_D}
\newcommand{\acontrans}{\stackrel{A}{\rightarrow_C}}
\newcommand{\adistrans}{\stackrel{A}{\rightarrow_D}}
\newcommand{\Dur}{{\IR_{\geq 0}}}
\newcommand{\leltdec}{{\tt leltdec}}
\newcommand{\leltdecoverlap}{{\tt leltdecoverlap}}
\newcommand{\leltdecDN}{{\tt leltdecDN}}

@@

\begin{abstract}
We have developed a hybrid system safety prover, implemented in Coq
using the abstraction method introduced by Alur, Dang and Ivan\v ci\'c
(2006).  The development includes: a formalisation of the structure of
hybrid systems; a framework for the construction of an abstract
system (consisting of decidable ``overestimators'' of abstract
transitions and initiality) faithfully representing a (concrete)
hybrid system; a translation of abstract systems to graphs, enabling
the decision of abstract state reachability using a certified graph
reachability algorithm; a proof of the safety of an example hybrid
system generated using this tool stack. 
To produce fully certified safety proofs without relying on floating point computations, the development critically relies on the computable real number implementation part of the CoRN library of constructive mathematics formalised in Coq.
The development also features a nice interplay between constructive and 
classical logic via the double negation monad.

@@

In \cite{alur}, Alur et al.\ present an automated method for hybrid system
safety verification in which one derives from the hybrid system of interest an \emph{abstract}
hybrid system, which is essentially a finite automaton whose traces are sufficiently representative of traces in the original system that unreachability in the abstract system (which can be decided using a standard graph algorithm) implies unreachability in the concrete system (which, being governed by continuous behaviors, cannot be decided so readily). Thus, the abstraction method brings the safety verification problem from a continuous and infinite domain into a discrete and finite domain, where it can be dealt with using standard graph algorithms.

@@

\section{Hybrid Systems and the Abstraction method}
\label{sec:hybsys}
In general, a hybrid system is a model of how a software system (running on a device with sensors and actuators), described as a finite set of \emph{locations}
%\footnote{We reserve the term ``state'' for something else.}
with (discrete) transitions between them, acts
on and responds to a set of continuous variables (called the
\emph{continuous state space}), typically representing physical
properties of some environment (such as temperature and pressure).

There are many varieties of hybrid systems \cite{henziger,lynchvaandrager}. We follow \cite{alur} in considering so-called \emph{linear hybrid systems}\weg{ (to be defined momentarily)}. To illustrate the definition and the abstraction process, we use as an example a hybrid system expressing the operation of a thermostat (this is the same example used in \cite{alur}), shown in Figure \ref{fig:thermostat}.

@@

With each location is associated an \emph{invariant} predicate defining the set of
permitted values for the continuous variables while in that location. The invariants for the thermostat are: $$\Inv_{\Heat} := c\leq 3 \wedge T\leq 10,\quad \Inv_{\Cool} := 5 \leq T,\quad \Inv_{\Chec} := c\leq 1.$$

@@

The \emph{initial states} of a hybrid system are determined by a predicate $\Init$. For the thermostat,
$\Init(l,c,T) := l= \Heat \wedge c= 0 \wedge 5 \leq T \leq 10.$

@@

The remaining parts of a hybrid system (described below) describe transitions between states, which, together
with the set of initial states, determine the set of \emph{reachable}
states, representing the possible behaviours exhibited by a
hypothetical real-world implementation of the hybrid system (as
software running on a device with sensors and actuators).

Each location in a hybrid system has an accompanying \emph{flow function} which describes how the continuous variables change over
time while the system is in that location. The idea is that different locations correspond to different uses of actuators
available to the software system, the effects of which are reflected in the flow function. For instance, in the thermostat, the flow
function corresponding to the $\Heat$ location will have the
temperature increase over time, modelling the effect of the heater
component in the imagined thermostat device.

In the canonical definition of hybrid systems, flow functions are
specified as solutions to differential equations describing the
dynamics of the continuous variables. We follow \cite{alur} in
abstracting from these, taking instead the solutions of these
differential equations, which are flow functions $\Phi$ which satisfy:
$$\Phi(p, 0) = p \quad \mbox{and} \quad \Phi (p, d + d') = \Phi (\Phi(p, d), d')$$
We now say that there is a (concrete) \emph{continuous transition}
from a state $(l, p)$ to a state $(l, p')$ if there is a
non-negative duration $d$ such that $p' = \Phi_l(p,d)$ with the
invariant for $l$ holding at every point along the way:
$$ (l,p) \contrans (l',p') := l = l' \wedge \Exists{d\in
\Dur}{\Phi_l(p, d) = p' \wedge \Forall{0 \leq t\leq d}{\Inv_l(\Phi_l (p, t))}}.$$

A flow function on |pow IR 2| can be expressed as the
product of two flow functions: $\Phi_l((c,T),d) = (\phi_{l,c}((c,T),d),
\phi_{l,T}((c,T),d))$. In the thermostat example, as in many other
examples of hybrid systems, $\phi_{l,c}((c,T),d)$ does not actually depend on $T$
and $\phi_{l,T}((c,T),d)$ does not actually depend on $c$. We call this feature
{\em separability\/} of the flow function. Our development currently
relies heavily on this property. Separability makes the form of the
flow functions simpler:
$$\Phi_l((c,T),t) = (\phi_{l,c}(c,t), \phi_{l,T}(T,t))$$

In the thermostat, $\phi_{l,c}(c,d) = c + d$ for all locations $l$, $\phi_{\Heat,T}(T,d)= T + 2 d$,
$\phi_{\\Cool,T}(T,d)= T * e^{-d}$, $\phi_{\Chec,T}(T,d)= T * e^{- \frac{1}{2}d}$.

Where continuous transitions describe the flow of continuous
variables, \emph{discrete} transitions between locations describe the
logic of the software system. Each such transition is comprised of two
components: a \emph{guard} predicate defining a subset of the continuous state
space in which the transition is enabled (permitted), and a \emph{reset} function
describing an
instantaneous change applied as a side effect of the transition, as
seen in the following definition of the discrete transition relation:
$$ (l,p) \distrans (l',p') := \guard_{l,l'}(p) \wedge \reset_{l,l'}(p)= p' \wedge \Inv_l(p) \wedge \Inv_{l'}(p')$$
It will be obvious from Figure \ref{fig:thermostat} what the guards
and reset functions are.

@@

A transition is either continuous or discrete: $\trans:= \distrans \cup \contrans$. We now say that a state $s$ is \emph{reachable} if there is an initial state $i$ from which one can, by a finite number of transitions, end up in $s$ ($\ttrans$ is the transitive reflexive closure of $\trans$):
$$\Reach(s):=
    \Exists{i\in \State}{\Init(i) \wedge i \ttrans s}.$$
\weg{As mentioned before, the set of reachable states represents the
possible behaviours exhibited by a hypothetical real-world
implementation of the hybrid system (as software running on a device
with sensors and actuators).
}

The object of hybrid system safety verification is to show that the set of reachable states is a subset of a predefined set of ``safe'' states. For the thermostat, the intent is to
keep the temperature above $4.5$ degrees at all times, and so we define $\Safe(c,T) := T > 4.5$ (and $\Unsafe(c,T)$ as its complement).

@@

The idea is to divide the continuous state space
into a finite number of convex subsets (polygons), $A_1, \ldots, A_n$,
which yields a finite {\em abstract state space}, $\AState := \{ (l,
A_i) \mid l \in \Loc, 1\leq i \leq n \}$, with an obvious embedding
$A: \State \to \AState$ of concrete states into abstract states. On this
abstract state space, one immediately defines {\em abstract continuous
  transitions\/} and {\em abstract discrete transitions\/} (both potentially undecidable) as follows.
\begin{eqnarray*}
(l,P) \acontrans (l',Q) &:=& l=l' \wedge \Exists{p\in P, q\in Q}{(l,p) \contrans (l,q)}\\
(l,P) \adistrans (l',Q) &:=& \Exists{p\in P, q\in Q}{(l,p) \distrans (l,q)}.
\end{eqnarray*}
Now {\em abstract reachability\/} can be defined in the obvious way as
follows $\AReach(a) := \Exists{s_0\in\State}{\Init(s_0) \wedge A(s_0)
\attrans a}$, and also the predicates $\ASafe$ and $\AUnsafe$, stating
when abstract states are safe / unsafe are defined as expected.

@@

The interest and power of the abstraction method lies in two
facts. First, we do not need the {\em exact\/} definitions of
$\acontrans$ and $\adistrans$ to conclude safety of the concrete
system from safety of the abstract system. We only need the property
of diagram \ref{diag:abstraction}, so we can {\em overestimate\/}
$\acontrans$ and $\adistrans$ (i.e. replace it with a transition
relation that allows more transitions). Second, there are good
heuristics for how to divide the continuous
state space into regions, and how to decide whether there should be an
abstract transition from one abstract state to another.

This is indicated in figure \ref{fig:abstraction}. The left side illustrates
the challenge: given abstract regions $A$ and $B$, we are to determine
whether nonnegative flow durations might permit flow from points in $A$ to points in $B$.
Following the overestimation property just mentioned, we introduce an abstract transition from $A$ to $B$ whenever we cannot positively rule out such paths.

On the right side we see the abstract state space indicated for the
location $\Heat$. The abstract state space consists of rectangles,
possibly degenerate (extending to $-\infty$ or $+\infty$). For the
bounds of these rectangles one takes the values that occur in the
specification of the hybrid system (see Figure
\ref{fig:thermostat}). According to \cite{alur}, this is a good
choice. (In case one cannot prove safety in this, there is of course
the opportunity for the user to refine it.)  The gray area indicates
that from these states also abstract discrete transitions are
possible. The dashed area is unreachable, because of the invariant for
the $\Heat$ location. %prevents these states from being reachable.  
All the abstract
transitions from the rectangle $[0.5,1) \times[5,6)$ are shown: as the
    temperature flow function for $\Heat$ is $\phi_{\Heat,2}(T,d)= T +
    2 * d$, and the clock flow function is $\phi_{\Heat,1}(c,d) = c +
    d$, these are all the possible abstract transitions.

@@

\begin{center}
\begin{figure}[htb!]
%\begin{turn}{180}
%    \includegraphics[scale=.6, viewport=0 500 700 710, clip]{FIGS/traject}
    \includegraphics[scale=.5]{FIGS/traject}
    \includegraphics[scale=.4]{FIGS/AThermo-new3}
%\end{turn}
\caption{The abstraction function computed
\label{fig:abstraction}}
\end{figure}
\end{center}
\weg{ % because this is explained by Alur already and it is not of particular relevance in our development; we don't refer to it ever again.
There is a slight subtlety with the abstract transitions that we want
to consider. If we just take $\attrans$ to be the reflexive transitive
closure of $\acontrans \cup \adistrans$, we get far too many abstract
traces. (There is no problem with the correctness of the method --
Figure \ref{diag:abstraction} still applies -- but we get way too many
traces to be able to prove any system safe.) The problem lies in the
fact that there is no transition from any point in $[0.5,1)
  \times[5,6)$ to $[0.5,1) \times[9,10)$, but the transitive closure
        of $\adistrans$ gives us such a transition anyway.
}
% \ADAM{I will try to add some more explanation}
\section{Formalisation}
We now describe the Coq formalisation and the design choices made.
We will not pay much attention to the specifics of Coq and its type theory CiC, and will instead focus on concerns relating to the use of computable reals and constructive logic.

The complete sources of the development, as well as a technical report describing the formalization in more detail \cite{hybrid-techreport}, can be found at \newline {\tt http://www.eelis.net/research/hybrid/}.

@@

\subsection{Abstract Hybrid Systems}
\label{abshybsys}
\label{abstraction}
We now want to define an abstract system and an abstraction function
satisfying the properties indicated in diagram
\ref{diag:abstraction}. However, this is not possible, because we
cannot make a case distinction like $x\leq 0 \vee 0< x$ and therefore
we can't define a function that sends a point $(c,T)$ to the rectangle $R$
it is in. We can define a function that approximates a point $(c,T)$
up to, say $\epsilon$ ($\epsilon >0$) and then decides to send that
point to the rectangle $R$ the approximation is in. This implies that,
when one is close to the edge of a rectangle,
\begin{itemize}
\item 
different {\em representations\/} of a point $(c,T)$ may be
sent to different rectangles,
\item
a point that is less than $\epsilon$ outside the rectangle $R$ may still be sent to $R$.
\end{itemize}
The second is very problematic, because it means the property for the
abstraction function $A$ depicted in diagram
\ref{diag:abstraction} no longer holds.

@@

The good thing is that we don't really need the property in Diagram
\ref{diag:abstraction}, because we can do something better. 
\begin{itemize}
\item
We let regions in the abstract hybrid systems overlap (ideally as
little as possible, e.g.\ only at the edges).

\item
We replace the abstract transition relations $\acontrans$ and $\distrans$ by functions $\overcontrans$ and $\overdistrans$ that take a region $R_0$
as input and output a {\em list of regions\/} including $R_0$: $(R_o,
R_1, \ldots, R_n)$ in such a way that $\cup_{0\leq i\leq n}R_i$ is an
overapproximation of the set of states reachable by a continuous (resp. discrete) step
from a state in $R_0$.\\
\item
We loosen the requirement on the abstraction function $A$; for $s \in \State$, we only require $\Varid{DN}(\Exists {r\in \Region}{s\in r})$.
\end{itemize}

To summarize, if $s\contrans s'$, then we don't require
$A(s')$ to be in the list $\overcontrans(A(s))$, but we only require
$s'$ to be in the $\bigcup \overcontrans(A(s))$. 
%(Basically it should
%be in one of the regions in the list $\overcontrans(A(s))$.) 
This simple change relieves us from having to determine the exact
regions that points are in: they just should be covered. The functions
$\overcontrans$ and $\overdistrans$ yield a notion of {\em trace\/} in
the abstract hybrid system in the obvious way: starting from $R_0$,
take an $R_1$ in $\overcontrans( R_0)$, then an $R_2$ in
$\overdistrans(R_1)$, and so forth.

@@

It is \emph{necessary} because |regions_cover| boils down to a
(partial) function that, given a concrete point, must select an
abstract region containing that point. This means that it must be able
to decide on which side of a border between two regions the given
point lies. As we saw in section \ref{dn}, that kind of decidability
is only available inside |DN| unless all region borders have
nontrivial overlap, which as we will see later is undesireable.

@@

\subsection{Underestimation and Overestimation}
\label{estimation}

% It \emph{will} always terminate, so we are not writing a semidecision procedure either.
Ultimately, in our development we are writing a program that
\emph{attempts} to produce hybrid system safety proofs. Importantly,
we are \emph{not} writing a complete hybrid system safety decision
procedure: if the concrete system is unsafe or the abstraction method
fails, our program will simply not produce a safety proof. It might
seem, then, that we are basically writing a \emph{tactic} for a
particular problem domain. However, tactics in Coq are normally
written in a language called Ltac, and typically rely on things like
pattern matching on syntax. Our development, on the other hand, is
very much written in regular Gallina, with hardly any significant use
of Ltac.

>>>>

\documentclass[runningheads]{llncs}
\ifx\pdftexversion\undefined
  \usepackage[dvips]{graphics}
\else
  \usepackage[pdftex]{graphics}
\fi
\usepackage{amsmath}
%include format.tex
%\usepackage{rotating}
\usepackage{graphicx}
\input{diagrams}
\newcommand{\thetitle}{Automated Machine-Checked Hybrid System Safety Proofs\footnote{This research was supported by the BRICKS/FOCUS project 642.000.501, Advancing the Real use of Proof Assistants}}
\newcommand{\eelis}{Eelis van der Weegen$^{\dagger}$}
\newcommand{\herman}{Herman Geuvers$^{\dagger\ddagger}$}
\newcommand{\dan}{Dan Synek$^{\dagger}$}
\newcommand{\adam}{Adam Koprowski$^{\P}$}

@@

\def\phi{\varphi}
\newcommand{\weg}[1]{}
\newcommand{\DN}{{\sf DN}\,}
\newcommand{\IR}{{\mathbf R}}
\newcommand{\ST}{{\sf T}}
\newcommand{\Sc}{{\sf c}}
\newcommand{\Heat}{{\sf Heat}}
\newcommand{\Cool}{{\sf Cool}}
\newcommand{\Chec}{{\sf Check}}
\newcommand{\Off}{{\sf Off}} 
\newcommand{\Loc}{{\rm Loc}}
\newcommand{\AState}{{\rm AState}}
\newcommand{\State}{{\rm State}}
\newcommand{\Inv}{{\rm Inv}}
\newcommand{\Init}{{\rm Init}}
\newcommand{\Region}{{\rm Region}}
\newcommand{\List}{{\rm List}}
\newcommand{\Unsafe}{{\rm Unsafe}}
\newcommand{\Safe}{{\rm Safe}}
\newcommand{\AUnsafe}{{\rm AUnsafe}}
\newcommand{\ASafe}{{\rm ASafe}}
\newcommand{\Reach}{{\rm Reach}}
\newcommand{\AReach}{{\rm AReach}}
\newcommand{\guard}{{\rm guard}}
\newcommand{\reset}{{\rm reset}}
\newcommand{\trans}{\rightarrow_{CD}}
\newcommand{\doubleheadrightarrow}{\mathrel{\rightarrow\!\!\!\!\rightarrow}}
\newcommand{\ttrans}{\doubleheadrightarrow_{CD}}
\newcommand{\attrans}{\stackrel{A}{\doubleheadrightarrow_{CD}}}
\newcommand{\contrans}{\rightarrow_C}
\newcommand{\distrans}{\rightarrow_D}
\newcommand{\acontrans}{\stackrel{A}{\rightarrow_C}}
\newcommand{\adistrans}{\stackrel{A}{\rightarrow_D}}
\newcommand{\Dur}{{\IR_{\geq 0}}}
\newcommand{\leltdec}{{\tt leltdec}}
\newcommand{\leltdecoverlap}{{\tt leltdecoverlap}}
\newcommand{\leltdecDN}{{\tt leltdecDN}}

@@

\begin{abstract}
We have developed a hybrid system safety prover, implemented in Coq
using the abstraction method introduced by Alur, Dang and Ivan\v ci\'c
(2006).  The development includes: a formalisation of the structure of
hybrid systems; a framework for the construction of an abstract
system (consisting of decidable ``over-estimators'' of abstract
transitions and initiality) faithfully representing a (concrete)
hybrid system; a translation of abstract systems to graphs, enabling
the decision of abstract state reachability using a certified graph
reachability algorithm; a proof of the safety of an example hybrid
system generated using this tool stack. 
To produce fully certified safety proofs without relying on floating point computations, the development critically relies on the computable real number implementation part of the CoRN library of constructive mathematics formalised in Coq.
The development also features a nice interplay between constructive and 
classical logic via the double negation monad.

@@

In \cite{alur}, Alur et al.\ present an automated method for hybrid system
safety verification in which one derives from the hybrid system of interest an \emph{abstract}
hybrid system, which is essentially a finite automaton whose traces are sufficiently representative of traces in the original system that unreachability in the abstract system (which can be decided using a standard graph algorithm) implies unreachability in the concrete system (which, being governed by continuous behaviours, cannot be decided so readily). Thus, the abstraction method brings the safety verification problem from a continuous and infinite domain into a discrete and finite domain, where it can be dealt with using standard graph algorithms.

@@

\section{Hybrid Systems and the Abstraction method}
\label{sec:hybsys}
A hybrid system is a model of how a software system (running on a device with sensors and actuators), described as a finite set of \emph{locations}
%\footnote{We reserve the term ``state'' for something else.}
with (discrete) transitions between them, acts
on and responds to a set of continuous variables (called the
\emph{continuous state space}), typically representing physical
properties of some environment (such as temperature and pressure).

There are many varieties of hybrid systems
\cite{henziger,lynchvaandrager}. We follow \cite{alur} in considering
so-called \emph{linear hybrid systems}\weg{ (to be defined
  momentarily)}. To illustrate the definition and the abstraction
process, we use the example of a hybrid system describing
a thermostat (this is the same example used in \cite{alur}), shown
in Figure \ref{fig:thermostat}. (The linearity in \cite{alur} refers
to invariants, guards etc.; our development doesn't use that.)

@@

With each location is associated an \emph{invariant} predicate defining the set of
permitted values for the continuous variables while in that location. The invariants for the thermostat are: $$\Inv_{\Heat}(c,T) :=  T\leq 10 \wedge c\leq 3,\quad \Inv_{\Cool}(c,T) := T \geq 5,\quad \Inv_{\Chec}(c,T) := c\leq 1.$$

@@

The \emph{initial states} of a hybrid system are determined by a predicate $\Init$. For the thermostat,
$\Init(l,c,T)$ is defined as $l= \Heat \wedge c= 0 \wedge 5 \leq T \leq 10.$

@@

%The remaining parts of a hybrid system (described below) describe
%transitions between states, which, together with the set of initial
%states, determine the set of \emph{reachable} states, representing the
%possible behaviours exhibited by a hypothetical real-world
%implementation of the hybrid system (as software running on a device
%with sensors and actuators).
The \emph{discrete} transitions between locations describe the
logic of the software system. Each such transition is comprised of two
components: a \emph{guard} predicate defining a subset of the continuous state
space in which the transition is enabled (permitted), and a \emph{reset} function
describing an
instantaneous change applied as a side effect of the transition, as
seen in the following definition of the discrete transition relation:
$$ (l,p) \distrans (l',p') := \guard_{l,l'}(p) \wedge
\reset_{l,l'}(p)= p' \wedge \Inv_l(p) \wedge \Inv_{l'}(p')$$ It will
be clear from Figure \ref{fig:thermostat} what the guards and reset
functions are. Note the inherent non-determinism in a Hybrid Systems
specification: when in $\Cool$, the system can jump to $\Heat$
whenever the temperature $T$ is in the interval $[5,6]$.

Each location in a hybrid system has an accompanying \emph{flow function} which describes how the continuous variables change over
time while the system is in that location. The idea is that different locations correspond to different uses of actuators
available to the software system, the effects of which are reflected in the flow function. In the thermostat example, the flow
function corresponding to the $\Cool$ location has the
temperature decrease over time. 
%modelling the effect of the heater
%component in the imagined thermostat device.
This is expressed via the differential equation $\dot{T} = - T$,
which is the usual short hand for $T'(t) = - T(t)$, where $T'(t)$
denotes the derivative of the temperature function over time $t$.

In the canonical definition of hybrid systems, flow functions are
specified as solutions to differential equations (or differential
inclusions) describing the dynamics of the continuous variables. We
follow \cite{alur} in abstracting from these, taking instead the
solutions of these differential equations, which are flow functions
$\Phi$ which satisfy:
$$\Phi(p, 0) = p \quad \mbox{and} \quad \Phi (p, d + d') = \Phi
(\Phi(p, d), d')$$ 
The idea is that $\Phi(p,d)$ denotes the value of the
continuous variable after duration $d$, starting from the value
$p$. We say that there is a (concrete) \emph{continuous
  transition} from a state $(l, p)$ to a state $(l, p')$ if there is a
non-negative duration $d$ such that $p' = \Phi_l(p,d)$ with the
invariant for $l$ holding at every point along the way:
$$ (l,p) \contrans (l,p') := \Exists{d\in
\Dur}{\Phi_l(p, d) = p' \wedge \Forall{0 \leq t\leq d}{\Inv_l(\Phi_l (p, t))}}.$$

A flow function on |pow IR 2| can be expressed as the
product of two flow functions: $\Phi_l((c_0,T_0),t) = (\phi_{l,\Sc}((c_0,T_0),t),
\phi_{l,\ST}((c_0,T_0),t))$. In the thermostat example, as in many other
examples of hybrid systems, $\phi_{l,\Sc}((c_0,T_0),t)$ does not actually depend on $T_0$
and $\phi_{l,\ST}((c_0,T_0),t)$ does not actually depend on $c_0$. We call this feature
{\em separability\/} of the flow function. Our development currently
relies heavily on this property. Separability makes the form of the
flow functions simpler:
$$\Phi_l((c_0,T_0),t) = (\phi_{l,\Sc}(c_0,t), \phi_{l,\ST}(T_0,t))$$

In the thermostat, $\phi_{l,\Sc}(c_0,t) = c_0 + t$ for all locations
$l$, $\phi_{\Heat,\ST}(T_0,t)= T_0 + 2 t$, $\phi_{\Chec,\ST}(T_0,t)=
T_0 * e^{- \frac{1}{2}t}$ and $\phi_{\Cool,\ST}(T_0,t)= T_0 *
e^{-t}$. So $\phi'_{\Cool,\ST}(T_0,t)= - \phi_{\Cool,\ST}(T_0,t)$,
solving the differential equation $\dot{T} = - T$ for the $\Cool$
location.

@@

A transition is either continuous or discrete: $\trans:= \distrans
\cup \contrans$. A finite sequence of transitions constitutes a {\em trace}. We now say that a state $s$ is \emph{reachable} if
there is an initial state $i$ from which there is a trace to $s$.
% one can, by a finite number
%of transitions, end up in $s$ (
If we denote by $\ttrans$ the transitive reflexive
closure of $\trans$, that is:
$$\Reach(s):=
    \Exists{i\in \State}{\Init(i) \wedge i \ttrans s}.$$

\weg{As mentioned before, the set of reachable states represents the
possible behaviours exhibited by a hypothetical real-world
implementation of the hybrid system (as software running on a device
with sensors and actuators).
}

The objective of hybrid system safety verification is to show that the set of reachable states is a subset of a predefined set of ``safe'' states. For the thermostat, the intent is to
keep the temperature above $4.5$ degrees at all times, and so we define $\Safe(c,T) := T > 4.5$ (and $\Unsafe(c,T)$ as its complement).

@@

The idea is to divide the continuous state space
into a finite number of convex subsets (polygons), $A_1, \ldots, A_n$,
which yields a finite {\em abstract state space}, $\AState := \{ (l,
A_i) \mid l \in \Loc, 1\leq i \leq n \}$, with an obvious embedding
$A: \State \to \AState$ of concrete states into abstract states. On this
abstract state space, one immediately defines {\em abstract continuous
  transitions\/} and {\em abstract discrete transitions\/} (both potentially undecidable) as follows.
\begin{eqnarray*}
(l,P) \acontrans (l,Q) &:=& \Exists{p\in P, q\in Q}{(l,p) \contrans (l,q)}\\
(l,P) \adistrans (l',Q) &:=& \Exists{p\in P, q\in Q}{(l,p) \distrans (l,q)}.
\end{eqnarray*}
Now {\em abstract reachability\/} can be defined as
follows $\AReach(a) := \Exists{s_0\in\State}{\Init(s_0) \wedge A(s_0)
\attrans a}$, and also the predicates $\ASafe$ and $\AUnsafe$, stating
when abstract states are safe / unsafe are defined as expected.

@@

The interest and power of the abstraction method lies in two
facts. First, we do not need the {\em exact\/} definitions of
$\acontrans$ and $\adistrans$ to conclude safety of the concrete
system from safety of the abstract system. We only need the property
of Figure \ref{diag:abstraction}, so we can {\em over-estimate\/}
$\acontrans$ and $\adistrans$ (i.e. replace it with a transition
relation that allows more transitions). Second, there are good
heuristics for how to divide the continuous
state space into regions, and how to decide whether there should be an
abstract transition from one abstract state to another.

This is indicated in figure \ref{fig:abstraction}. The left hand side
illustrates the challenge: given abstract regions $A$ and $B$, we are
to determine whether some flow duration permits flow from points in
$A$ to points in $B$.  Following the over-estimation property just
mentioned, we introduce an abstract transition from $A$ to $B$
whenever we cannot positively rule this out.

On the right hand side we see the abstract state space indicated for the
location $\Heat$. The abstract state space consists of rectangles,
possibly degenerate (extending to $-\infty$ or $+\infty$). For the
bounds of these rectangles one takes the values that occur in the
specification of the hybrid system (see Figure
\ref{fig:thermostat}). According to \cite{alur}, this is a good
choice. (In case one cannot prove safety, there is of course
the opportunity for the user to refine the bounds.)  The grey area indicates
that from these states also abstract discrete transitions are
possible. The dashed area is unreachable, because of the invariant for
the $\Heat$ location. %prevents these states from being reachable.  
All the abstract
transitions from the rectangle $[0.5,1) \times[5,6)$ are shown: as the
    temperature flow function for $\Heat$ is $\phi_{\Heat,\ST}(T_0,t)= T_0 +
    2 * t$, and the clock flow function is $\phi_{\Heat,\Sc}(c_0,t) = c_0 +
    t$, these are all the possible abstract transitions.

@@

\begin{center}
\begin{figure}[htb!]
%\begin{turn}{180}
%    \includegraphics[scale=.6, viewport=0 500 700 710, clip]{FIGS/traject}
    \includegraphics[scale=.5]{FIGS/traject}
    \includegraphics[scale=.4]{FIGS/AThermo-new3}
%\end{turn}
\caption{The abstraction function computed
\label{fig:abstraction}}
\end{figure}
\end{center}
\weg{ % because this is explained by Alur already and it is not of particular relevance in our development; we don't refer to it ever again.
There is a slight subtlety with the abstract transitions that we want
to consider. If we just take $\attrans$ to be the reflexive transitive
closure of $\acontrans \cup \adistrans$, we get far too many abstract
traces. (There is no problem with the correctness of the method --
Figure \ref{diag:abstraction} still applies -- but we get way too many
traces to be able to prove any system safe.) The problem lies in the
fact that there is no transition from any point in $[0.5,1)
  \times[5,6)$ to $[0.5,1) \times[9,10)$, but the transitive closure
        of $\adistrans$ gives us such a transition anyway.
}
% \ADAM{I will try to add some more explanation}
\section{Formalisation}
We now describe the Coq formalisation and the design choices made.  We
will not pay much attention to the specifics of Coq and its type
theory CiC, and will instead focus on concerns relating to the use of
computable reals and constructive logic.  The complete sources of the
development are available on the web, as well as a technical report
describing the formalisation in more detail \cite{hybrid-techreport}.
%, can be found
%at \newline {\tt http://www.eelis.net/research/hybrid/}.

@@

\subsection{Abstract Hybrid Systems}
\label{abshybsys}
\label{abstraction}
We now want to define an abstract system and an abstraction function
satisfying the properties indicated in Figure
\ref{diag:abstraction}. However, this is not possible, because we
cannot make a case distinction like $x\leq 0 \vee 0< x$ and therefore
we can't define a function that sends a point $(c,T)$ to the rectangle $R$
it is in. We can define a function that approximates a point $(c,T)$
up to, say $\epsilon$ ($\epsilon >0$) and then decides to send that
point to the rectangle $R$ the approximation is in. This implies that,
when one is close to the edge of a rectangle,
\begin{itemize}
\item 
different {\em representations\/} of a point $(c,T)$ may be
sent to different rectangles,
\item
a point that is less than $\epsilon$ outside the rectangle $R$ may still be sent to $R$.
\end{itemize}
The second is very problematic, because it means the property for the
abstraction function $A$ depicted in Figure
\ref{diag:abstraction} no longer holds.

@@

The good thing is that we don't really need the property in Figure
\ref{diag:abstraction}, because we can do something better. 
\begin{itemize}
\item
We let regions in the abstract hybrid systems overlap (ideally as
little as possible, e.g.\ only at the edges).

\item
We replace the abstract relations $\acontrans$ and $\adistrans$ by functions $\overcontrans$ and $\overdistrans$ that take a region $R_0$
as input and output a {\em list of regions\/} including $R_0$: $(R_o,
R_1, \ldots, R_n)$ in such a way that $\cup_{0\leq i\leq n}R_i$ is an
over-approximation of the set of states reachable by a continuous (resp. discrete) step
from a state in $R_0$.\\
\item
We loosen the requirement on the abstraction function $A$; for $s \in \State$, we only require $\Varid{DN}(\Exists {r\in \Region}{s\in r})$.
\end{itemize}

To summarise, if $s\contrans s'$, then we don't require
$A(s')$ to be in the list $\overcontrans(A(s))$, but we only require
$s'$ to be in the $\bigcup \overcontrans(A(s))$. 
%(Basically it should
%be in one of the regions in the list $\overcontrans(A(s))$.) 
This simple change relieves us from having to determine the exact
regions that points are in: they just should be covered. The functions
$\overcontrans$ and $\overdistrans$ yield a notion of {\em trace\/} in
the abstract hybrid system in the straightforward way: starting from $R_0$,
take an $R_1$ in $\overcontrans( R_0)$, then an $R_2$ in
$\overdistrans(R_1)$, and so forth.

@@

It is \emph{necessary} because |regions_cover| boils down to a
(partial) function that, given a concrete point, must select an
abstract region containing that point. This means that it must be able
to decide on which side of a border between two regions the given
point lies. As we saw in section \ref{dn}, that kind of decidability
is only available inside |DN| unless all region borders have
nontrivial overlap, which 
%as we will see later 
is undesirable.

@@

\subsection{Under-Estimation and Over-Estimation}
\label{estimation}

% It \emph{will} always terminate, so we are not writing a semi-decision procedure either.
Ultimately, in our development we are writing a program that
\emph{attempts} to produce hybrid system safety proofs. Importantly,
we are \emph{not} writing a complete hybrid system safety decision
procedure: if the concrete system is unsafe or the abstraction method
fails, our program will simply not produce a safety proof. It might
seem, then, that we are basically writing a \emph{tactic} for a
particular problem domain. However, tactics in Coq are normally
written in a language called Ltac, and typically rely on things like
pattern matching on syntax. Our development, on the other hand, is
very much written in regular Gallina, with hardly any significant use
of Ltac.
